The Bolocam Galactic Plane Survey: Survey Description and Data Reduction

### 2009-04-17 methods/ms.tex
remove rms from plot header
introduction modified with Schuller materials

\newcommand{\ammonia}{\mbox{{\rm NH}$_3$}}
\newcommand{\form}{H$_2$CO}
\newcommand{\td}{\mbox{$T_D$}}
\newcommand{\jj}[2]{\mbox{$J = #1\rightarrow#2$}}
\newcommand{\vdag}{(v)^\dagger}
\newcommand{\myemail}{{\tt jaguirre@sas.upenn.edu}}
\newcommand{\lsim}{{_{<}\atop^{\sim}}}
\newcommand{\gsim}{{_{>}\atop^{\sim}}}
\newcommand{\etal}{{et al.\/}}
\newcommand{\ie}{{\em ie.\/}}
\newcommand{\cmq}{cm{$^{-3}$}}
\newcommand{\per}{$^{\rm{-1}}$}
\newcommand{\tc}{{$\theta^1$~Orionis~C}}
%\newcommand{\msol}{M{$_{\odot}$}}
\def\msol{\ifmmode {\>M_\odot}\else {$M_\odot$}\fi}
\newcommand{\lsol}{L{$_{\odot}$}}
\newcommand{\kms}{km~s{$^{-1}$}}
\newcommand{\hii}{H~{\sc ii}}
\newcommand{\Hii}{H~{\sc ii}}
\newcommand{\Ha}{\mbox{H$\alpha$}}
\newcommand{\sii}{S~{\sc ii}}
\newcommand{\Feii}{Fe~{\sc ii}}
\newcommand{\oi}{O~{\sc i}}
\newcommand{\nii}{N~{\sc ii}}
\newcommand{\oiii}{O~{\sc iii}}
\newcommand{\mgii}{Mg~{\sc ii}}
\newcommand{\tco}{{$^{13}$CO}}
\newcommand{\CO}{{$^{12}$CO}}
\newcommand{\Tco}{{$^{12}$CO}}
\newcommand{\co}{C{$^{18}$O}}
\newcommand{\Lsol}{L$_{\odot}$}
\newcommand{\Msol}{M$_{\odot}$}
%\newcommand{\C18o}{C$^{18}$O($1\rightarrow 0$)}
\newcommand{\Check}{{\bf ???}}
\newcommand{\mum}{\ensuremath{\mu \mathrm{m}}}
\newcommand{\flux}{flux density}
\newcommand{\solar}{\ensuremath{\odot}}
\newcommand{\lon}{\ensuremath{\ell}}

@@

To pursue this question, we need tracers of higher density material.
Studies of nearby clouds have demonstrated that star formation is far
from uniform over molecular clouds, but is concentrated in unusually dense
regions called clumps or cores, which fill a small fraction of the area
of most clouds [refs. E. Lada et al. Evans et al., Enoch et al.]
Following \citet{williams00} and McKee-Ostriker (ARAA), we
use the term `core' to refer to a very dense region destined to form an
individual star or small multiple star systems and the term `clump' to
refer to a
region likely to form a group or cluster of stars. While clumps forming
massive clusters may be denser than cores forming low mass stars
[ref Mueller?],
in general, clumps are larger and less dense. Clumps can be traced
with line emission from molecules such as \ammonia, \form, CS, HCN, etc.
Spectral lines also provide excellent diagnostics of
line-of-sight motions, temperatures, and densities.
However, the interpretation of these gas tracers is complicated by
variations in tracer abundances caused by depletions and complex
chemical processing, uncertainties in optical depths and excitation
conditions, and the impacts of radiation fields and shocks. These
complications make the derivation of column densities, masses, and
other physical properties of star-forming clumps very difficult.
A further limitation of spectral line studies is that the tracers of
denser gas tend to be weak and multi-element receiver arrays have
only a modest number of elements, making blind surveys very time-consuming.

% redundant:
% In contrast, focal-plane arrays containing hundreds of individual
% bolometers sensitive to millimeter and sub-millimeter (sub-mm)
% radiation are now available. This advance in technology enables
% blind surveys of the Galactic plane, providing a uniform inventory of massive
% star-forming and starless clumps.  Catalogs resulting
% from such surveys will provide the data base for subsequent observations
% in tracers such as hard-to-excite molecular lines and higher resolution
% studies with instruments like CARMA and ALMA.
(this line belongs somewhere)
A further advantage of continuum observations of dust at long wavelengths
is the low optical depth of the dust \citep{johnstone06},
permitting a straight-forward estimate of the mass of the emitting region.

@@

Two large surveys of the Galactic plane are in progress.
The ATLASGAL survey, using the LABOCA instrument at 870 \micron\
on the APEX telescope in Chile, will cover the southern Galactic plane
and extend somewhat into the north. Initial results from this survey
have been reported by \citet{schuller09}. [say more - overlap region?
]

@@

% 9
% -----------------------------------------------------------------------------
\begin{figure}
  \subfigure[The pointing model correction (left) and residuals
(right) for Epoch V, from which the master reference images were
derived for subsequent alignment.  The red line indicates the fitted
model. ]
{\label{fig:PointingModel-a}\includegraphics[scale=0.9]{paper_models_0707}}\\

>>>>

\newcommand{\ammonia}{\mbox{{\rm NH}$_3$}}
\newcommand{\form}{H$_2$CO}
\newcommand{\td}{\mbox{$T_D$}}
\newcommand{\jj}[2]{\mbox{$J = #1\rightarrow#2$}}
\newcommand{\vdag}{(v)^\dagger}
\newcommand{\myemail}{{\tt jaguirre@sas.upenn.edu}}
\newcommand{\lsim}{{_{<}\atop^{\sim}}}
\newcommand{\gsim}{{_{>}\atop^{\sim}}}
\newcommand{\etal}{{et al.\/}}
\newcommand{\ie}{{\em ie.\/}}
\newcommand{\cmq}{cm{$^{-3}$}}
\newcommand{\per}{$^{\rm{-1}}$}
\newcommand{\tc}{{$\theta^1$~Orionis~C}}
%\newcommand{\msol}{M{$_{\odot}$}}
\def\msol{\ifmmode {\>M_\odot}\else {$M_\odot$}\fi}
\newcommand{\lsol}{L{$_{\odot}$}}
\newcommand{\kms}{km~s{$^{-1}$}}
\newcommand{\hii}{H~{\sc ii}}
\newcommand{\Hii}{H~{\sc ii}}
\newcommand{\Ha}{\mbox{H$\alpha$}}
\newcommand{\sii}{S~{\sc ii}}
\newcommand{\Feii}{Fe~{\sc ii}}
\newcommand{\oi}{O~{\sc i}}
\newcommand{\nii}{N~{\sc ii}}
\newcommand{\oiii}{O~{\sc iii}}
\newcommand{\mgii}{Mg~{\sc ii}}
\newcommand{\tco}{{$^{13}$CO}}
\newcommand{\CO}{{$^{12}$CO}}
\newcommand{\Tco}{{$^{12}$CO}}
\newcommand{\co}{C{$^{18}$O}}
\newcommand{\Lsol}{L$_{\odot}$}
\newcommand{\Msol}{M$_{\odot}$}
%\newcommand{\C18o}{C$^{18}$O($1\rightarrow 0$)}
\newcommand{\Check}{{\bf ???}}
\newcommand{\mum}{\ensuremath{\mu \mathrm{m}}}
\newcommand{\flux}{flux density}
\newcommand{\solar}{\ensuremath{\odot}}
\newcommand{\lon}{\ensuremath{l}}

@@

To pursue this question, we need tracers of higher density material.
Studies of nearby clouds have demonstrated that star formation is far
from uniform over molecular clouds, but is concentrated in unusually dense
regions called clumps or cores, which fill a small fraction of the area of most
clouds \citep{enoch06,evans09}[refs. E. Lada et al. Evans et al., Enoch et al.]
(Neal, John, Jonathan) Following \citet{williams00} and \citep{mckee07}, we use
the term `core' to refer to a very dense region destined to form an individual
star or small multiple star systems and the term `clump' to refer to a region
likely to form a group or cluster of stars. While clumps forming massive
clusters may be denser than cores forming low mass stars [ref Mueller? (Neal,
John, Jonathan)], in general, clumps are larger and less dense. Clumps can be
traced with line emission from molecules such as \ammonia, \form, CS, HCN, etc.
Spectral lines also provide excellent diagnostics of line-of-sight motions,
temperatures, and densities.  However, the interpretation of these gas tracers
is complicated by variations in tracer abundances caused by depletions and
complex chemical processing, uncertainties in optical depths and excitation
conditions, and the impacts of radiation fields and shocks. These complications
make the derivation of column densities, masses, and other physical properties
of star-forming clumps very difficult.  A limitation of spectral line studies
is that the tracers of denser gas tend to be weak and multi-element receiver
arrays have only a modest number of elements, making blind surveys very
time-consuming.

In contrast, focal-plane arrays containing hundreds of individual bolometers
sensitive to millimeter and sub-millimeter (sub-mm) radiation are now
available. This advance in technology enables blind surveys of the Galactic
plane, providing a uniform inventory of massive star-forming and starless
clumps.  Catalogs resulting from these surveys will provide the data base for
subsequent observations in tracers such as hard-to-excite molecular lines and
higher resolution studies with instruments like CARMA and ALMA.  A significant
advantage of continuum observations of dust at long wavelengths is the low
optical depth of the dust \citep{johnstone06}, permitting a straight-forward
estimate of the mass of the emitting region.

@@

Two large surveys of the Galactic plane are in progress.  The ATLASGAL survey,
using the LABOCA instrument at 870 \micron\  with 12\arcsec resolution on the
APEX telescope in Chile, will cover the southern Galactic plane and some of the
north.  Initial results from this survey have been reported by
\citet{schuller09}. The overlap region between BGPS and ATLASGAL is in the
region $-10.5 < \lon < 11.5$ and $15 < \lon < 21$, $-.5 < b < .5$, though
ATLASGAL will eventually expand up to $\lon = 60$.  The two surveys provide
complementary samples of the long-wavelength end of the spectral energy
distribution in Galactic star forming regions.

@@

% 9
% -----------------------------------------------------------------------------
\begin{figure}
  \subfigure[The pointing model correction (left) and residuals (right) for
  Epoch V, from which the master reference images were derived for subsequent
  alignment.  The red line indicates the fitted model. 510 pointing sources
  were included, and the final RMS was 5.77\arcsec in altitude and 3.21\arcsec
  in azimuth, or a total RMS offset of 6.6\arcsec]
{\label{fig:PointingModel-a}\includegraphics[scale=0.9]{paper_models_0707}}\\
